---
title: "US insulin estimates"
author: "Sanjay Basu"
date: "3/3/2019"
output: html_document
---

```{r setup, include=FALSE}
# install.packages("knitr")
# install.packages("dplyr")
# install.packages("nhanesA")
# install.packages("fitdistrplus")
# install.packages("readxl")
# install.packages("MatchIt")
# install.packages("mice")
# install.packages("tableone")
# install.packages("captioner")
# install.packages("survey")
# install.packages("data.table")
# install.packages("formattable")
# install.packages("stargazer")
# install.packages("ggplot2")
# install.packages("plyr")
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warnings = FALSE)
#setwd("~/Box/Research/Research projects/Insulin/US insulin needs")
library(dplyr)
library(nhanesA)
library(fitdistrplus)
library(readxl)
library(readr)
library(MatchIt)
library(mice)
library(tableone)
library(captioner)
library(survey)
library(data.table)
library(formattable)
library(stargazer)
library(ggplot2)
library(plyr)


```
## 4t analysis of units per weight

```{r 4t}
JYdataset01 <- read_csv("JY4TData/JYdataset01.csv")
JYdataset02 <- read_csv("JY4TData/JYdataset02.csv")
fourt= JYdataset01 %>%
  left_join(JYdataset02,by=c("Patient_ID" = "Patient_ID"))
summary(as.numeric(fourt$ins1)/fourt$weight0)
fourt$ins1 = as.numeric(fourt$ins1)
fourt$ins0 = as.numeric(fourt$ins0)
fourt$a1c1 = as.numeric(fourt$a1c1)
fourt$a1c0 = as.numeric(fourt$a1c0)
fourt$a1cd = fourt$a1c1-fourt$a1c0
fourt$met1 = as.numeric(fourt$met1)
fourt$sulf1 = as.numeric(fourt$sulf1)
fourt$insd = fourt$ins1-fourt$ins0
fourt$race = as.factor(fourt$race)

reg1  = lm(a1cd~weight0+ins1+age+female+as.factor(race)+met1+sulf1, data=fourt)
summary(reg1)

star.1 <- stargazer(reg1,
                    title="Multivariable regression from the 4-T Study",
                    align=TRUE,
                    type = "text",
                    style = "default",
                    ci=TRUE,
                    ci.level=0.95,
                    single.row=T
                    )
knitr::kable(star.1)


```


## NHANES download and organize


```{r nhanes assemble}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warnings = FALSE)

demo_i <- nhanes('DEMO_I')
diq_i <- nhanes('DIQ_I')
ghb_i <- nhanes('GHB_I')
glu_i <- nhanes('GLU_I')
bpx_i <- nhanes('BPX_I')
bpq_i <- nhanes('BPQ_I')
mcq_i <- nhanes('MCQ_I')
hdl_i <- nhanes('HDL_I')
tchol_i <- nhanes('TCHOL_I')
biopro_i <- nhanes('BIOPRO_I')
albcr_i <- nhanes('ALB_CR_I')
smq_i <- nhanes('SMQ_I')
bmx_i <- nhanes('BMX_I')
hiq_i <- nhanes('HIQ_I')


demo_h <- nhanes('DEMO_H')
diq_h <- nhanes('DIQ_H')
ghb_h <- nhanes('GHB_H')
glu_h <- nhanes('GLU_H')
bpx_h <- nhanes('BPX_H')
bpq_h <- nhanes('BPQ_H')
mcq_h <- nhanes('MCQ_H')
hdl_h <- nhanes('HDL_H')
tchol_h <- nhanes('TCHOL_H')
biopro_h <- nhanes('BIOPRO_H')
albcr_h <- nhanes('ALB_CR_H')
smq_h <- nhanes('SMQ_H')
bmx_h <- nhanes('BMX_H')
hiq_h <- nhanes('HIQ_H')


demo_g <- nhanes('DEMO_G')
diq_g <- nhanes('DIQ_G')
ghb_g <- nhanes('GHB_G')
glu_g <- nhanes('GLU_G')
bpx_g <- nhanes('BPX_G')
bpq_g <- nhanes('BPQ_G')
mcq_g <- nhanes('MCQ_G')
hdl_g <- nhanes('HDL_G')
tchol_g <- nhanes('TCHOL_G')
biopro_g <- nhanes('BIOPRO_G')
albcr_g <- nhanes('ALB_CR_G')
smq_g <- nhanes('SMQ_G')
bmx_g <- nhanes('BMX_G')
hiq_g <- nhanes('HIQ_G')


demo_f <- nhanes('DEMO_F')
diq_f <- nhanes('DIQ_F')
ghb_f <- nhanes('GHB_F')
glu_f <- nhanes('GLU_F')
bpx_f <- nhanes('BPX_F')
bpq_f <- nhanes('BPQ_F')
mcq_f <- nhanes('MCQ_F')
hdl_f <- nhanes('HDL_F')
tchol_f <- nhanes('TCHOL_F')
biopro_f <- nhanes('BIOPRO_F')
albcr_f <- nhanes('ALB_CR_F')
smq_f <- nhanes('SMQ_F')
bmx_f <- nhanes('BMX_F')
hiq_f <- nhanes('HIQ_F')


demo_e <- nhanes('DEMO_E')
diq_e <- nhanes('DIQ_E')
ghb_e <- nhanes('GHB_E')
glu_e <- nhanes('GLU_E')
bpx_e <- nhanes('BPX_E')
bpq_e <- nhanes('BPQ_E')
mcq_e <- nhanes('MCQ_E')
hdl_e <- nhanes('HDL_E')
tchol_e <- nhanes('TCHOL_E')
biopro_e <- nhanes('BIOPRO_E')
albcr_e <- nhanes('ALB_CR_E')
smq_e <- nhanes('SMQ_E')
bmx_e <- nhanes('BMX_E')
hiq_e <- nhanes('HIQ_E')


demo_d <- nhanes('DEMO_D')
diq_d <- nhanes('DIQ_D')
ghb_d <- nhanes('GHB_D')
glu_d <- nhanes('GLU_D')
bpx_d <- nhanes('BPX_D')
bpq_d <- nhanes('BPQ_D')
mcq_d <- nhanes('MCQ_D')
hdl_d <- nhanes('HDL_D')
tchol_d <- nhanes('TCHOL_D')
biopro_d <- nhanes('BIOPRO_D')
albcr_d <- nhanes('ALB_CR_D')
smq_d <- nhanes('SMQ_D')
bmx_d <- nhanes('BMX_D')
hiq_d <- nhanes('HIQ_D')



nhanes1516 <- left_join(demo_i, diq_i, by = c('SEQN'))
nhanes1516 <- left_join(nhanes1516,ghb_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,glu_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,bpx_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,bpq_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,mcq_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,hdl_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,tchol_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,biopro_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,albcr_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,smq_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,bmx_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,hiq_i, by = c('SEQN'))

nhanes1314 <- left_join(demo_h, diq_h, by = c('SEQN'))
nhanes1314 <- left_join(nhanes1314,ghb_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,glu_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,bpx_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,bpq_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,mcq_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,hdl_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,tchol_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,biopro_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,albcr_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,smq_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,bmx_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,hiq_h, by = c('SEQN'))

nhanes1112 <- left_join(demo_g, diq_g, by = c('SEQN'))
nhanes1112 <- left_join(nhanes1112,ghb_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,glu_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,bpx_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,bpq_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,mcq_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,hdl_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,tchol_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,biopro_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,albcr_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,smq_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,bmx_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,hiq_g, by = c('SEQN'))

nhanes0910 <- left_join(demo_f, diq_f, by = c('SEQN'))
nhanes0910 <- left_join(nhanes0910,ghb_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,glu_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,bpx_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,bpq_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,mcq_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,hdl_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,tchol_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,biopro_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,albcr_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,smq_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,bmx_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,hiq_f, by = c('SEQN'))

nhanes0708 <- left_join(demo_e, diq_e, by = c('SEQN'))
nhanes0708 <- left_join(nhanes0708,ghb_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,glu_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,bpx_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,bpq_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,mcq_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,hdl_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,tchol_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,biopro_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,albcr_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,smq_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,bmx_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,hiq_e, by = c('SEQN'))

nhanes0506 <- left_join(demo_d, diq_d, by = c('SEQN'))
nhanes0506 <- left_join(nhanes0506,ghb_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,glu_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,bpx_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,bpq_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,mcq_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,hdl_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,tchol_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,biopro_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,albcr_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,smq_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,bmx_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,hiq_d, by = c('SEQN'))


nhanes0506  = nhanes0506 %>%
  mutate(year = 2005)
nhanes0708  = nhanes0708 %>%
  mutate(year = 2007)
nhanes0910  = nhanes0910 %>%
  mutate(year = 2009)
nhanes1112 = nhanes1112 %>%
  mutate(year = 2011)
nhanes1314 = nhanes1314 %>%
  mutate(year = 2013)
nhanes1516 = nhanes1516 %>%
  mutate(year = 2015)

nhanes1516 <- nhanes1516 %>%
  dplyr::select("RIDAGEYR","RIAGENDR","RIDRETH1", "BMXWT","URDACT","SMQ040","BPXSY1","BPQ050A","LBDHDD","LBXTC","LBXSCR", "MCQ160E","MCQ160F", "DIQ010","LBXGH","WTSAF2YR","LBXGLU","DIQ050","DIQ070","LBXSATSI","DID040", "INDFMPIR", "WTINT2YR", "SDMVPSU", "SDMVSTRA","BMXWAIST", "HIQ031B", "HIQ031D","INDFMPIR") %>%
  as_tibble()

nhanes1314 <- nhanes1314 %>%
  dplyr::select("RIDAGEYR","RIAGENDR","RIDRETH1", "BMXWT","URDACT","SMQ040","BPXSY1","BPQ050A","LBDHDD","LBXTC","LBXSCR", "MCQ160E","MCQ160F", "DIQ010","LBXGH","WTSAF2YR","LBXGLU","DIQ050","DIQ070","LBXSATSI","DID040", "INDFMPIR", "WTINT2YR", "SDMVPSU", "SDMVSTRA","BMXWAIST", "HIQ031B", "HIQ031D","INDFMPIR") %>%
  as_tibble()


nhanes1112 <- nhanes1112 %>%
  dplyr::select("RIDAGEYR","RIAGENDR","RIDRETH1", "BMXWT","URDACT","SMQ040","BPXSY1","BPQ050A","LBDHDD","LBXTC","LBXSCR", "MCQ160E","MCQ160F", "DIQ010","LBXGH","WTSAF2YR","LBXGLU","DIQ050","DIQ070","LBXSATSI","DID040", "INDFMPIR", "WTINT2YR", "SDMVPSU", "SDMVSTRA","BMXWAIST", "HIQ031B", "HIQ031D","INDFMPIR") %>%
  as_tibble()

nhanes0910 <- nhanes0910 %>%
  dplyr::select("RIDAGEYR","RIAGENDR","RIDRETH1", "BMXWT","URDACT","SMQ040","BPXSY1","BPQ050A","LBDHDD","LBXTC","LBXSCR", "MCQ160E","MCQ160F", "DIQ010","LBXGH","WTSAF2YR","LBXGLU","DIQ050","DIQ070","LBXSATSI","DID040", "INDFMPIR", "WTINT2YR", "SDMVPSU", "SDMVSTRA","BMXWAIST", "HIQ031B", "HIQ031D","INDFMPIR") %>%
  as_tibble()

nhanes0708 <- nhanes0708 %>%
  mutate(DIQ070 = DID070,
         URDACT = URXUMS/URXUCR) %>%
  dplyr::select("RIDAGEYR","RIAGENDR","RIDRETH1", "BMXWT","URDACT","SMQ040","BPXSY1","BPQ050A","LBDHDD","LBXTC","LBXSCR", "MCQ160E","MCQ160F", "DIQ010","LBXGH","WTSAF2YR","LBXGLU","DIQ050","DIQ070","LBXSATSI","DID040", "INDFMPIR", "WTINT2YR", "SDMVPSU", "SDMVSTRA","BMXWAIST", "HIQ031B", "HIQ031D","INDFMPIR") %>%
  as_tibble() 

nhanes0506 <- nhanes0506  %>%
  mutate(DIQ070 = DID070,
         URDACT = URXUMS/URXUCR) %>%
  dplyr::select("RIDAGEYR","RIAGENDR","RIDRETH1", "BMXWT","URDACT","SMQ040","BPXSY1","BPQ050A","LBDHDD","LBXTC","LBXSCR", "MCQ160E","MCQ160F", "DIQ010","LBXGH","WTSAF2YR","LBXGLU","DIQ050","DIQ070","LBXSATSI","DID040", "INDFMPIR", "WTINT2YR", "SDMVPSU", "SDMVSTRA","BMXWAIST", "HIQ031B", "HIQ031D","INDFMPIR") %>%
  as_tibble() 


nhanes <- rbind(nhanes1314,nhanes1516)
nhanes <- rbind(nhanes,nhanes1112)
nhanes <- rbind(nhanes,nhanes0910)
nhanes <- rbind(nhanes,nhanes0708)
nhanes <- rbind(nhanes,nhanes0506)



# age, female, black, sbp, bprx, oraldmrx, cvd_hx_baseline,hba1c,chol,hdl,screat,uacr
#rm(nhanes0910,nhanes1112,nhanes1314,nhanes1516,demo_g,demo_h,demo_i,demo_f,diq_g,diq_h,diq_i,diq_f,ghb_g,ghb_h,ghb_i,ghb_f,glu_g,glu_h,glu_i,glu_f,bpx_g,bpx_h,bpx_i,bpx_f)
save(nhanes,file="nhanes")


```

## NHANES A1c distributions

```{r nhanes a1c distribs}


load("nhanes")

dfn = nhanes[(complete.cases(nhanes$WTSAF2YR)&complete.cases(nhanes$LBXGH)),]

dfn$persWeight <- dfn$WTSAF2YR/6
dfn$psu        <- dfn$SDMVPSU
dfn$strata     <- dfn$SDMVSTRA


dfn$dm = (dfn$DIQ010==1) | 
  (dfn$LBXGH>=6.5) | 
  (dfn$LBXGLU>=126)
dfn <- dfn[which(dfn$dm==1), ]
table(dfn$dm)

dfn$dx = dfn$DIQ010
dfn$oral = dfn$DIQ070
dfn$ins = dfn$DIQ050
dfn$oral[is.na(dfn$DIQ070)]=2
dfn$ins[is.na(dfn$DIQ050)]=2
dfn$dx[dfn$dx==3]=2
dfn$dx[dfn$dx==9]=2
dfn$oral[dfn$oral==9]=2
dfn$ins[dfn$ins==9]=2

dfn$grp = 1*(dfn$dx==2)+#undiagnosed
  2*(dfn$dx==1)*(dfn$oral==2)*(dfn$ins==2)+#diagnosed and untreated
  3*(dfn$dx==1)*(dfn$oral==1)*(dfn$ins==2)+#diagnosed and non-insulin rx only
  4*(dfn$dx==1)*(dfn$ins==1) #diagnosed and insulin rx


dfn$agedx = dfn$DID040
#dfn$agedx[dfn$agedx>80]='NA'

dfn$durdx = dfn$RIDAGEYR-as.numeric(dfn$agedx)
dfn$durdx[dfn$durdx<0]=0

dfnDesign <- svydesign(id      = ~psu,
                          strata  = ~strata,
                          weights = ~persWeight,
                          nest    = TRUE,
                          data    = dfn)

# TABLE 1
table(dfn$grp)
table(dfn$grp)/sum(table(dfn$grp))
svytable(~grp, dfnDesign)
svytable(~grp, dfnDesign)/sum(svytable(~grp, dfnDesign))
svyby(~RIDAGEYR,~grp,dfnDesign, svymean)
svyby(~RIDAGEYR,~grp,dfnDesign, svyquantile,quantiles=c(.25,.75),keep.var=F)
svytable(~RIAGENDR+grp,dfnDesign)/as.data.frame(svytable(~grp, dfnDesign))$Freq
svytable(~RIDRETH1+grp, dfnDesign)/as.data.frame(svytable(~grp, dfnDesign))$Freq
svyby(~durdx,~grp,dfnDesign, svymean, na.rm=T)
#svyby(~durdx,~grp,dfnDesign, svyquantile,quantiles=c(.25,.75),keep.var=F, na.rm=T)
svyby(~LBXGH,~grp,dfnDesign, svymean, na.rm=T)
svyby(~LBXGH,~grp,dfnDesign, svyquantile,quantiles=c(.25,.75),keep.var=F, na.rm=T)
svyby(~LBXGLU,~grp,dfnDesign, svymean, na.rm=T)
svyby(~LBXGLU,~grp,dfnDesign, svyquantile,quantiles=c(.25,.75),keep.var=F, na.rm=T)
svyby(~BMXWT,~grp,dfnDesign, svymean, na.rm=T)
svyby(~BMXWT,~grp,dfnDesign, svyquantile,quantiles=c(.25,.75),keep.var=F,na.rm=T)


# TABLE 2
dfn$basea1c =  1*(dfn$dx==2)*(dfn$LBXGH>7)+#undiagnosed
  2*(dfn$dx==1)*(dfn$oral==2)*(dfn$ins==2)*(dfn$LBXGH>7)+#diagnosed and untreated
  3*(dfn$dx==1)*(dfn$oral==1)*(dfn$ins==2)*(dfn$LBXGH>7)+#diagnosed and non-insulin rx only
  4*(dfn$dx==1)*(dfn$ins==1)*(dfn$LBXGH>7) #diagnosed and insulin rx

dfn$grpa1c = 1*(dfn$dx==2)*(dfn$LBXGH-2.5>7)+#undiagnosed
  2*(dfn$dx==1)*(dfn$oral==2)*(dfn$ins==2)*((dfn$LBXGH-2.5)>7)+#diagnosed and untreated
  3*(dfn$dx==1)*(dfn$oral==1)*(dfn$ins==2)*((dfn$LBXGH-(2.5/2))>7)+#diagnosed and non-insulin rx only
  4*(dfn$dx==1)*(dfn$ins==1)*((dfn$LBXGH-(2.5/2)*(table(dfn$oral,dfn$ins)[1,2]/sum(table(dfn$oral,dfn$ins))))>7) #diagnosed and insulin rx

dfn$newa1coral = (dfn$grpa1c==1)*(dfn$LBXGH-2.5)+
  (dfn$grpa1c==2)*(dfn$LBXGH-2.5)+
  (dfn$grpa1c==3)*(dfn$LBXGH-(2.5/2))+
  (dfn$grpa1c==4)*(dfn$LBXGH-((2.5/2)*(table(dfn$oral,dfn$ins)[1,2]/sum(table(dfn$oral,dfn$ins)))))
  
dfn$delta = (dfn$newa1coral-7)
dfn$delta[dfn$delta<0] = 0

dfn$insneed = dfn$BMXWT*0.6419

dfnDesign <- svydesign(id      = ~psu,
                          strata  = ~strata,
                          weights = ~persWeight,
                          nest    = TRUE,
                          data    = dfn)


sum(dfn$basea1c>0)
table(dfn$basea1c)
table(dfn$basea1c)/sum(dfn$basea1c>0)
svytable(~basea1c, dfnDesign)
svytable(~basea1c, dfnDesign)/(svytable(~basea1c>0, dfnDesign))[2]
svyby(~LBXGH,~basea1c>0,dfnDesign, svymean, na.rm=T)
svyby(~LBXGH,~basea1c,dfnDesign, svyquantile,quantiles=c(.25,.75),keep.var=F)

sum(dfn$grpa1c>0)
table(dfn$grpa1c)
table(dfn$grpa1c)/sum(dfn$grpa1c>0)
svytable(~grpa1c, dfnDesign)
svytable(~grpa1c, dfnDesign)/(svytable(~grpa1c>0, dfnDesign))[2]
svyby(~newa1coral,~grpa1c,dfnDesign, svymean, na.rm=T)
svyby(~newa1coral,~grpa1c,dfnDesign, svyquantile,quantiles=c(.25,.75),keep.var=F)
svyby(~BMXWT,~grpa1c,dfnDesign, svymean, na.rm=T)
svyby(~BMXWT,~grpa1c,dfnDesign, svyquantile,quantiles=c(.25,.75),keep.var=F)
quantile(dfn$BMXWT[dfn$grpa1c==2],quantiles=c(.25,.75),na.rm=T)
quantile(dfn$BMXWT[dfn$grpa1c==4],quantiles=c(.25,.75),na.rm=T)
svyby(~insneed,~grpa1c,dfnDesign, svymean, na.rm=T)
svyby(~insneed,~grpa1c,dfnDesign, svyquantile,quantiles=c(.25,.75),keep.var=F)
quantile(dfn$insneed[dfn$grpa1c==2],quantiles=c(.25,.75),na.rm=T)
quantile(dfn$insneed[dfn$grpa1c==4],quantiles=c(.25,.75),na.rm=T)

svyby(~BMXWT,~grpa1c>0,dfnDesign, svymean, na.rm=T)
svyby(~BMXWT,~grpa1c>0,dfnDesign,svyquantile,quantiles=c(.25,.75),keep.var=F,na.rm=T)

svyby(~insneed,~grpa1c>0,dfnDesign, svymean, na.rm=T)
svyby(~insneed,~grpa1c>0,dfnDesign,svyquantile,quantiles=c(.25,.75),keep.var=F,na.rm=T)

```

# Insulin secular trend

```{r secular trend in insulin use}
nhanes0506$dm = (nhanes0506$DIQ010==1) | 
  (nhanes0506$LBXGH>=6.5) | 
  (nhanes0506$LBXGLU>=126)

nhanes0708$dm = (nhanes0708$DIQ010==1) | 
  (nhanes0708$LBXGH>=6.5) | 
  (nhanes0708$LBXGLU>=126)

nhanes0910$dm = (nhanes0910$DIQ010==1) | 
  (nhanes0910$LBXGH>=6.5) | 
  (nhanes0910$LBXGLU>=126)

nhanes1112$dm = (nhanes1112$DIQ010==1) | 
  (nhanes1112$LBXGH>=6.5) | 
  (nhanes1112$LBXGLU>=126)

nhanes1314$dm = (nhanes1314$DIQ010==1) | 
  (nhanes1314$LBXGH>=6.5) | 
  (nhanes1314$LBXGLU>=126)

nhanes1516$dm = (nhanes1516$DIQ010==1) | 
  (nhanes1516$LBXGH>=6.5) | 
  (nhanes1516$LBXGLU>=126)

table(nhanes0506$DIQ050==1,nhanes0506$dm==1)[2,2]/sum(nhanes0506$dm==1,na.rm=T)
table(nhanes0708$DIQ050==1,nhanes0708$dm==1)[2,2]/sum(nhanes0708$dm==1,na.rm=T)
table(nhanes0910$DIQ050==1,nhanes0910$dm==1)[2,2]/sum(nhanes0910$dm==1,na.rm=T)
table(nhanes1112$DIQ050==1,nhanes1112$dm==1)[2,2]/sum(nhanes1112$dm==1,na.rm=T)
table(nhanes1314$DIQ050==1,nhanes1314$dm==1)[2,2]/sum(nhanes1314$dm==1,na.rm=T)
table(nhanes1516$DIQ050==1,nhanes1516$dm==1)[2,2]/sum(nhanes1516$dm==1,na.rm=T)

```

# Demographic change approach: 

Demographic change approach: multiply number of people with diagnosed diabetes from 2020-2030 (from IDF estimates) by proportion of people anticipated to be treated with insulin given current estimates (from NHANES) by amount of insulin per person (given distribution of body weights in NHANES).


```{r approach 1}


# https://diabetesatlas.org/data/en/country/211/us.html
undx = .381
total_dm_19 = 30987.90*1000
total_dm_30 = 34374.20*1000

type2prop = 0.93
type2proplo = 0.9
type2prophi = 0.95

proponins = 0.216

total_dm = total_dm_19 + (total_dm_30-total_dm_19)/(30-19)*(1:11)

# number of people with diagnosed type 2 dm
  n_dm1=(1-undx)*total_dm*type2prop
  n_dmlo1=(1-undx)*total_dm*type2proplo
  n_dmhi1=(1-undx)*total_dm*type2prophi

# 4t study regression
  
unitsperday_1 = (0.6419)*svymean(~BMXWT,dfnDesign,na.rm=T)[1]
unitsperdaylo_1 = (0.6419)*svyquantile(~BMXWT,dfnDesign,quantiles=c(.025),na.rm=T)[1]
unitsperdayhi_1 = (0.6419)*svyquantile(~BMXWT,dfnDesign,quantiles=c(.975),na.rm=T)[1]


# total vials across entire population of people with diagnosed type 2 dm per year
  vialsperyear_1=proponins*unitsperday_1*n_dm1*365.25/1000
  vialsperyearlo_1=proponins*unitsperdaylo_1*n_dmlo1*365.25/1000
  vialsperyearhi_1=proponins*unitsperdayhi_1*n_dmhi1*365.25/1000




```



# Demographic change plus improved insulin access: 
calculate additional number of people with diabetes not yet on insulin who would be diagnosed if US reached highest OECD country diagnosis level (UK, at 81.5% diagnosis per IDF Diabetes Atlas 9th edition). Estimate insulin necessary given current ADA-EASD guidelines (i.e. start with metformin, then either metformin+ SU, +TZD, +DPP-4, +SGLT-2, or +GLP-1RA). The base case should target HbA1c of 7%. [Sensitivity analyses not yet done: (i) For increased age or significant comorbidities, loosen target to 8%. (ii) try three orals before insulin. ]

```{r approach 2}


undx= 1-0.815


  n_dm2=(1-undx)*total_dm*type2prop
  n_dmlo2=(1-undx)*total_dm*type2proplo
  n_dmhi2=(1-undx)*total_dm*type2prophi


unitsperday_2 = (0.6419)*svymean(~BMXWT,dfnDesign,na.rm=T)[1]
unitsperdaylo_2 = (0.6419)*svyquantile(~BMXWT,dfnDesign,quantiles=c(.025),na.rm=T)[1]
unitsperdayhi_2 = (0.6419)*svyquantile(~BMXWT,dfnDesign,quantiles=c(.975),na.rm=T)[1]

# total units across entire population per year
  vialsperyear_2=unitsperday_2*n_dm2*365.25/1000*proponins
  vialsperyearlo_2=unitsperdaylo_2*n_dmlo2*365.25/1000*proponins
  vialsperyearhi_2=unitsperdayhi_2*n_dmhi2*365.25/1000*proponins

# table demo only  
yr = 2020:2030
dfneed=data.table::data.table(yr,
                           total_dm*type2prop,
                           n_dm1,
                           n_dm2,
                           n_dm1*proponins,
                           n_dm2*proponins,
                           vialsperyear_1,
                           vialsperyear_2)

dfneedlo=data.table::data.table(yr,
                           total_dm*type2proplo,
                           n_dmlo1,
                           n_dmlo2,
                           n_dmlo1*proponins,
                           n_dmlo2*proponins,
                           vialsperyearlo_1,
                           vialsperyearlo_2)

dfneedhi=data.table::data.table(yr,
                                total_dm*type2prophi,
                                n_dmhi1,
                                n_dmhi2,
                                n_dmhi1*proponins,
                                n_dmhi2*proponins,
                                vialsperyearhi_1,
                                vialsperyearhi_2)

formattable(dfneed)
formattable(dfneedlo)
formattable(dfneedhi)

```


#  Match NHANES People to MEPS to get imputed insulin dispensed

```{r match}

meps <- read_csv("mepsdiab021220.csv")


# meps$days = meps$TOTDAYS
# meps$days[is.na(meps$days)] = 1
meps$units = meps$TOTDOSE
# meps$units[is.na(meps$units)] = 0
meps$unitsperday = (meps$units/365)
meps$unitsperday[meps$unitsperday<=0] = 'NA'
meps$unitsperday[meps$unitsperday=='Inf'] = 'NA'

meps$unitsperday[meps$cover_whole_yr!=1] = 'NA'


# meps$DIABAGED[meps$DIABAGED<0]=0

mat_nhanes = dfn %>%
  mutate(baseline_age = RIDAGEYR,
         female = (RIAGENDR==2),
         black = (RIDRETH1==4),
         hisp = (RIDRETH1<3),
         tob = (SMQ040<3),
         sbp = BPXSY1,
         bprx = (BPQ050A==1),
        oraldmrx = (DIQ070==1),
        anti_coag = 0,
        cvd_hx_baseline = ((MCQ160E==1)|(MCQ160F==1)),
        hba1c = LBXGH,
        chol = LBXTC,
        hdl = LBDHDD,
        screat = LBXSCR,
        uacr = URDACT,
        alt = LBXSATSI,
        wt = BMXWT,
        yrsdiab =  (baseline_age - DID040),
        unitsperday = 0,
        postoral  = newa1coral,
        wc = BMXWAIST,
        mcare = (HIQ031B==15),
        mcaid = (HIQ031D==17),
        pir = INDFMPIR) %>%
  dplyr::select(baseline_age,
                female,
                black,
                hisp,
                tob,
                sbp,
                bprx,
                oraldmrx,
                anti_coag,
                cvd_hx_baseline,
                hba1c,
                chol,
                hdl,
                screat,
                uacr,
                alt,
                wt,
                yrsdiab,
                unitsperday,
                postoral,
                wc,
                mcare,
                mcaid,
                pir)

mat_nhanes$yrsdiab[mat_nhanes$yrsdiab<0] = 'NA'
mat_nhanes$yrsdiab = as.numeric(mat_nhanes$yrsdiab)
mat_nhanes$group = 1
mat_nhanes_imp = mice(mat_nhanes,m=1,maxit=10,meth='pmm',seed=1)
mat_nhanes_comp = complete(mat_nhanes_imp)

meps$DIABAGED = as.numeric(meps$DIABAGED)
meps$DIABAGED[meps$DIABAGED<0] = 0

mat_meps = meps %>%
  mutate(baseline_age = age,
         female = (SEX==2),
         black = (racex==2),
         hisp = 0,
         tob = 0,
         sbp =0,
         bprx = 0,
          oraldmrx = 0,
          anti_coag = 0,
          cvd_hx_baseline = 0,
          hba1c = 0,
          chol = 0, 
          hdl = 0,
          screat = 0,
          uacr = 0,
          alt = 0,
          wt = 0,
         yrsdiab = age - DIABAGED,
         postoral=0,
        wc = 0,
        mcare = 0,
        mcaid = 0,
        pir = 0) %>%
  dplyr::select(baseline_age,
                female,
                black,
                hisp,
                tob,
                sbp,
                bprx,
                oraldmrx,
                anti_coag,
                cvd_hx_baseline,
                hba1c,
                chol,
                hdl,
                screat,
                uacr,
                alt,
                wt,
                yrsdiab,
                unitsperday,
                postoral,
                wc,
                mcare,
                mcaid,
                pir)
mat_meps$yrsdiab[mat_meps$yrsdiab<0]=0
mat_meps$group = 0

mat_meps$unitsperday = as.numeric(mat_meps$unitsperday)

mat_meps_imp = mice(mat_meps,m=1,maxit=10,meth='pmm',seed=1)
mat_meps_comp = complete(mat_meps_imp)

df = rbind(mat_nhanes_comp,mat_meps_comp)

set.seed(1)
df$mcare[is.na(df$mcare)]=0
df$mcaid[is.na(df$mcaid)]=0

match.it <- matchit(group ~ baseline_age + female + black + yrsdiab, data = df, method="nearest", replace = T, caliper = 0.01, ratio = 1)
a <- summary(match.it)
matcheddf <- match.data(match.it)

table1 <- CreateTableOne(vars = c('baseline_age', 'female', 'black', 'yrsdiab'), 
                         data = df, 
                         factorVars = c('female','black'), 
                         strata = 'group')
table1 <- print(table1, 
                printToggle = FALSE, 
                noSpaces = TRUE, smd = TRUE)
knitr::kable(table1[,c(1,2,3,5)],  
      align = 'c', 
      caption = 'Table 1: Comparison of unmatched samples')

knitr::kable(a$nn, digits = 2, align = 'c', 
      caption = 'Table 2: Sample sizes')

table_match1 <- CreateTableOne(vars = c('baseline_age', 'female', 'black', 'yrsdiab'),strata = "group",data = matcheddf,test = T)
table3 <- print(table_match1, 
                printToggle = FALSE, 
                noSpaces = TRUE, smd = TRUE)
knitr::kable(table3,  
      align = 'c', 
      caption = 'Table 3: Summary of balance for matched data')



```


#   Calculate difference between estimated insulin needs (2020-2030) and imputed (based on current 2018 utilization).


```{r diff}

pairsdf = cbind(df[row.names(match.it$match.matrix),"wt"],
                df[match.it$match.matrix,"unitsperday"], 
                df[row.names(match.it$match.matrix),"hba1c"],
                df[row.names(match.it$match.matrix),"baseline_age"],
                df[row.names(match.it$match.matrix),"female"],
                df[row.names(match.it$match.matrix),"black"],
                df[row.names(match.it$match.matrix),"postoral"],
                df[row.names(match.it$match.matrix),"screat"],
                df[row.names(match.it$match.matrix),"alt"],
                df[row.names(match.it$match.matrix),"yrsdiab"],
                df[row.names(match.it$match.matrix),"hisp"],
                df[row.names(match.it$match.matrix),"pir"],
                df[row.names(match.it$match.matrix),"wc"],
                df[row.names(match.it$match.matrix),"mcare"],
                df[row.names(match.it$match.matrix),"mcaid"])

# pairsdf[,2][pairsdf[,2]>(mean(pairsdf[,2],na.rm=T)+sd(pairsdf[,2],na.rm=T))] = mean(pairsdf[,2])
# pairsdf[,1][pairsdf[,1]>(mean(pairsdf[,1],na.rm=T)+sd(pairsdf[,1],na.rm=T))] = mean(pairsdf[,1])
# pairsdf[,2][pairsdf[,2]<(mean(pairsdf[,2],na.rm=T)-sd(pairsdf[,2],na.rm=T))] = mean(pairsdf[,2])
# pairsdf[,1][pairsdf[,1]<(mean(pairsdf[,1],na.rm=T)-sd(pairsdf[,1],na.rm=T))] = mean(pairsdf[,1])

pairdfsub = pairsdf[pairsdf[,3]>7,]

a1c_before_orals = pairdfsub[,3]
# dfn$DIQ070[is.na(dfn$DIQ070)] = 9
# dfn$DIQ050[is.na(dfn$DIQ050)] = 9
# dfn$DIQ010[is.na(dfn$DIQ010)] = 9


a1c_after_orals = pairdfsub[,7]
a1c_after_orals[a1c_after_orals==0] = a1c_before_orals[a1c_after_orals==0]

wt = pairdfsub[,1]

age = pairdfsub[,4]
fem = pairdfsub[,5]
blk = pairdfsub[,6]
scr = pairdfsub[,8]
alt = pairdfsub[,9]
yrsdiab = pairdfsub[,10]
hisp = pairdfsub[,11]
pir = pairdfsub[,12]
wc = pairdfsub[,13]
mcare = pairdfsub[,14]
mcaid = pairdfsub[,15]

egfr = 175 * ((scr)^(-1.154)) * ((age)^(-0.203)) * (0.742*(fem==1)+1*(fem==0)) * (1.212 *(blk==1)+1*(blk==0))
ckd = egfr<60
cld = alt>56
  
need = (0.6419)*wt
summary(need)

used = pairdfsub[,2]

summary(used)
diff_disp_minus_need = used - need


summary(diff_disp_minus_need)
quantile(diff_disp_minus_need,c(.025,.975),na.rm=T)

# proportion over versus under treated

overrx = (used>(need+2*sd(need)))&(need>0)
underrx = (used<(need-2*sd(need)))&(need>0)
           
table(overrx==1)
table(underrx==1)

sum(underrx==1)/length(diff_disp_minus_need)
summary(diff_disp_minus_need[underrx==1])
quantile(diff_disp_minus_need[underrx==1],c(.025,.975),na.rm=T)


sum(overrx==1)/length(diff_disp_minus_need)
summary(diff_disp_minus_need[overrx==1])
quantile(diff_disp_minus_need[overrx==1],c(.025,.975),na.rm=T)


summary(need[underrx==1])
quantile(need[underrx==1],c(.025,.975),na.rm=T)

summary(need[overrx==1])
quantile(need[overrx==1],c(.025,.975),na.rm=T)


summary(used[underrx==1])
quantile(used[underrx==1],c(.025,.975),na.rm=T)

summary(used[overrx==1])
quantile(used[overrx==1],c(.025,.975),na.rm=T)


# a1c distribution in each group
summary(a1c_before_orals[underrx==1])
quantile(a1c_before_orals[underrx==1],c(.025,.975),na.rm=T)

summary(a1c_before_orals[overrx==1])
quantile(a1c_before_orals[overrx==1],c(.025,.975),na.rm=T)


summary(a1c_after_orals[underrx==1])
quantile(a1c_after_orals[underrx==1],c(.025,.975),na.rm=T)

summary(a1c_after_orals[overrx==1])
quantile(a1c_after_orals[overrx==1],c(.025,.975),na.rm=T)

summary(wt[underrx==1])
quantile(wt[underrx==1],c(.025,.975),na.rm=T)

summary(wt[overrx==1])
quantile(wt[overrx==1],c(.025,.975),na.rm=T)


summary(age[underrx==1])
quantile(age[underrx==1],c(.025,.975),na.rm=T)

summary(age[overrx==1])
quantile(age[overrx==1],c(.025,.975),na.rm=T)


summary(yrsdiab[underrx==1])
quantile(yrsdiab[underrx==1],c(.025,.975),na.rm=T)

summary(yrsdiab[overrx==1])
quantile(yrsdiab[overrx==1],c(.025,.975),na.rm=T)


summary(pir[underrx==1])
quantile(pir[underrx==1],c(.025,.975),na.rm=T)

summary(pir[overrx==1])
quantile(pir[overrx==1],c(.025,.975),na.rm=T)


summary(wc[underrx==1])
quantile(wc[underrx==1],c(.025,.975),na.rm=T)

summary(wc[overrx==1])
quantile(wc[overrx==1],c(.025,.975),na.rm=T)



table(fem[underrx==1])
table(fem[underrx==1])/length(fem[underrx==1])

table(fem[overrx==1])
table(fem[overrx==1])/length(fem[overrx==1])

table(blk[underrx==1])
table(blk[underrx==1])/length(blk[underrx==1])

table(blk[overrx==1])
table(blk[overrx==1])/length(blk[overrx==1])

table(hisp[underrx==1])
table(hisp[underrx==1])/length(hisp[underrx==1])

table(hisp[overrx==1])
table(hisp[overrx==1])/length(hisp[overrx==1])

table(mcare[underrx==1])
table(mcare[underrx==1])/length(mcare[underrx==1])

table(mcare[overrx==1])
table(mcare[overrx==1])/length(mcare[overrx==1])


table(mcaid[underrx==1])
table(mcaid[underrx==1])/length(mcaid[underrx==1])

table(mcaid[overrx==1])
table(mcaid[overrx==1])/length(mcaid[overrx==1])


table(ckd[underrx==1])
table(ckd[underrx==1])/length(ckd[underrx==1])

table(ckd[overrx==1])
table(ckd[overrx==1])/length(ckd[overrx==1])



table(cld[underrx==1])
table(cld[underrx==1])/length(cld[underrx==1])

table(cld[overrx==1])
table(cld[overrx==1])/length(cld[overrx==1])


save.image("USinsulin")
```

#   Hist


```{r hist}

df = data.frame(diff = diff_disp_minus_need,used=used,need=need)
p<-ggplot(df, aes(x=diff)) +
  geom_histogram(color="dodgerblue", fill="white",position="identity")
mu = ddply(df,"need",summarise, grp.sd=sd(need))
p+ scale_color_grey() + theme_classic()+
  labs(title="Difference between insulin dispensed and estimated need",x="Dispensed minus estimated need (IU/person/day)", y = "Count")+
  geom_vline(data=mu, aes(xintercept=-30.8, color='gray'),
             linetype="dashed")+
  geom_vline(data=mu, aes(xintercept=30.8, color='gray'),
             linetype="dashed")+
  theme(legend.position="none")


```

