---
title: "US insulin estimates"
author: "Sanjay Basu"
date: "3/3/2019"
output: html_document
---

```{r setup, include=FALSE}
# install.packages("knitr")
# install.packages("dplyr")
# install.packages("nhanesA")
# install.packages("fitdistrplus")
# install.packages("readxl")
# install.packages("MatchIt")
# install.packages("mice")
# install.packages("tableone")
# install.packages("captioner")
# install.packages("survey")
 
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warnings = FALSE)

library(dplyr)
library(nhanesA)
library(fitdistrplus)
library(readxl)
library(readr)
library(MatchIt)
library(mice)
library(tableone)
library(captioner)
library(survey)

```

## NHANES download and organize


```{r nhanes assemble}
demo_i <- nhanes('DEMO_I')
diq_i <- nhanes('DIQ_I')
ghb_i <- nhanes('GHB_I')
glu_i <- nhanes('GLU_I')
bpx_i <- nhanes('BPX_I')
bpq_i <- nhanes('BPQ_I')
mcq_i <- nhanes('MCQ_I')
hdl_i <- nhanes('HDL_I')
tchol_i <- nhanes('TCHOL_I')
biopro_i <- nhanes('BIOPRO_I')
albcr_i <- nhanes('ALB_CR_I')
smq_i <- nhanes('SMQ_I')
bmx_i <- nhanes('BMX_I')


demo_h <- nhanes('DEMO_H')
diq_h <- nhanes('DIQ_H')
ghb_h <- nhanes('GHB_H')
glu_h <- nhanes('GLU_H')
bpx_h <- nhanes('BPX_H')
bpq_h <- nhanes('BPQ_H')
mcq_h <- nhanes('MCQ_H')
hdl_h <- nhanes('HDL_H')
tchol_h <- nhanes('TCHOL_H')
biopro_h <- nhanes('BIOPRO_H')
albcr_h <- nhanes('ALB_CR_H')
smq_h <- nhanes('SMQ_H')
bmx_h <- nhanes('BMX_H')


demo_g <- nhanes('DEMO_G')
diq_g <- nhanes('DIQ_G')
ghb_g <- nhanes('GHB_G')
glu_g <- nhanes('GLU_G')
bpx_g <- nhanes('BPX_G')
bpq_g <- nhanes('BPQ_G')
mcq_g <- nhanes('MCQ_G')
hdl_g <- nhanes('HDL_G')
tchol_g <- nhanes('TCHOL_G')
biopro_g <- nhanes('BIOPRO_G')
albcr_g <- nhanes('ALB_CR_G')
smq_g <- nhanes('SMQ_G')
bmx_g <- nhanes('BMX_G')


demo_f <- nhanes('DEMO_F')
diq_f <- nhanes('DIQ_F')
ghb_f <- nhanes('GHB_F')
glu_f <- nhanes('GLU_F')
bpx_f <- nhanes('BPX_F')
bpq_f <- nhanes('BPQ_F')
mcq_f <- nhanes('MCQ_F')
hdl_f <- nhanes('HDL_F')
tchol_f <- nhanes('TCHOL_F')
biopro_f <- nhanes('BIOPRO_F')
albcr_f <- nhanes('ALB_CR_F')
smq_f <- nhanes('SMQ_F')
bmx_f <- nhanes('BMX_F')


demo_e <- nhanes('DEMO_E')
diq_e <- nhanes('DIQ_E')
ghb_e <- nhanes('GHB_E')
glu_e <- nhanes('GLU_E')
bpx_e <- nhanes('BPX_E')
bpq_e <- nhanes('BPQ_E')
mcq_e <- nhanes('MCQ_E')
hdl_e <- nhanes('HDL_E')
tchol_e <- nhanes('TCHOL_E')
biopro_e <- nhanes('BIOPRO_E')
albcr_e <- nhanes('ALB_CR_E')
smq_e <- nhanes('SMQ_E')
bmx_e <- nhanes('BMX_E')


demo_d <- nhanes('DEMO_D')
diq_d <- nhanes('DIQ_D')
ghb_d <- nhanes('GHB_D')
glu_d <- nhanes('GLU_D')
bpx_d <- nhanes('BPX_D')
bpq_d <- nhanes('BPQ_D')
mcq_d <- nhanes('MCQ_D')
hdl_d <- nhanes('HDL_D')
tchol_d <- nhanes('TCHOL_D')
biopro_d <- nhanes('BIOPRO_D')
albcr_d <- nhanes('ALB_CR_D')
smq_d <- nhanes('SMQ_D')
bmx_d <- nhanes('BMX_D')



nhanes1516 <- left_join(demo_i, diq_i, by = c('SEQN'))
nhanes1516 <- left_join(nhanes1516,ghb_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,glu_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,bpx_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,bpq_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,mcq_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,hdl_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,tchol_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,biopro_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,albcr_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,smq_i, by = c('SEQN'))
nhanes1516 <-   left_join(nhanes1516,bmx_i, by = c('SEQN'))

nhanes1314 <- left_join(demo_h, diq_h, by = c('SEQN'))
nhanes1314 <- left_join(nhanes1314,ghb_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,glu_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,bpx_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,bpq_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,mcq_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,hdl_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,tchol_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,biopro_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,albcr_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,smq_h, by = c('SEQN'))
nhanes1314 <-   left_join(nhanes1314,bmx_h, by = c('SEQN'))

nhanes1112 <- left_join(demo_g, diq_g, by = c('SEQN'))
nhanes1112 <- left_join(nhanes1112,ghb_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,glu_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,bpx_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,bpq_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,mcq_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,hdl_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,tchol_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,biopro_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,albcr_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,smq_g, by = c('SEQN'))
nhanes1112 <-  left_join(nhanes1112,bmx_g, by = c('SEQN'))

nhanes0910 <- left_join(demo_f, diq_f, by = c('SEQN'))
nhanes0910 <- left_join(nhanes0910,ghb_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,glu_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,bpx_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,bpq_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,mcq_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,hdl_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,tchol_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,biopro_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,albcr_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,smq_f, by = c('SEQN'))
nhanes0910 <-  left_join(nhanes0910,bmx_f, by = c('SEQN'))

nhanes0708 <- left_join(demo_e, diq_e, by = c('SEQN'))
nhanes0708 <- left_join(nhanes0708,ghb_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,glu_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,bpx_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,bpq_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,mcq_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,hdl_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,tchol_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,biopro_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,albcr_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,smq_e, by = c('SEQN'))
nhanes0708 <-  left_join(nhanes0708,bmx_e, by = c('SEQN'))

nhanes0506 <- left_join(demo_d, diq_d, by = c('SEQN'))
nhanes0506 <- left_join(nhanes0506,ghb_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,glu_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,bpx_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,bpq_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,mcq_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,hdl_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,tchol_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,biopro_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,albcr_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,smq_d, by = c('SEQN'))
nhanes0506 <-  left_join(nhanes0506,bmx_d, by = c('SEQN'))


nhanes0506  = nhanes0506 %>%
  mutate(year = 2005)
nhanes0708  = nhanes0708 %>%
  mutate(year = 2007)
nhanes0910  = nhanes0910 %>%
  mutate(year = 2009)
nhanes1112 = nhanes1112 %>%
  mutate(year = 2011)
nhanes1314 = nhanes1314 %>%
  mutate(year = 2013)
nhanes1516 = nhanes1516 %>%
  mutate(year = 2015)
nhanes <- bind_rows(nhanes1314,nhanes1516)
nhanes <- bind_rows(nhanes,nhanes1112)
nhanes <- bind_rows(nhanes,nhanes0910)
nhanes <- bind_rows(nhanes,nhanes0708)
nhanes <- bind_rows(nhanes,nhanes0506)

nhanes <- nhanes %>%
  dplyr::select("RIDAGEYR","RIAGENDR","RIDRETH1", "BMXWT","URDACT","SMQ040","BPXSY1","BPQ050A","LBDHDD","LBXTC","LBXSCR", "MCQ160E","MCQ160F", "DIQ010","LBXGH","WTSAF2YR","LBXGLU","DIQ050","DIQ070","LBXSATSI","DID040") %>%
  as_tibble()

# age, female, black, sbp, bprx, oraldmrx, cvd_hx_baseline,hba1c,chol,hdl,screat,uacr
rm(nhanes0910,nhanes1112,nhanes1314,nhanes1516,demo_g,demo_h,demo_i,demo_f,diq_g,diq_h,diq_i,diq_f,ghb_g,ghb_h,ghb_i,ghb_f,glu_g,glu_h,glu_i,glu_f,bpx_g,bpx_h,bpx_i,bpx_f)
save(nhanes,file="nhanes")

```

## NHANES A1c distributions

```{r nhanes a1c distribs}
load("nhanes")
nhanes$dm = (nhanes$DIQ010==1) | 
  (nhanes$LBXGH>=6.5) | 
  (nhanes$LBXGLU>=126)
nhanes <- nhanes[which(nhanes$dm==1), ]
nhanescomp = nhanes[(complete.cases(nhanes$WTSAF2YR)&complete.cases(nhanes$LBXGH)),]
table(nhanescomp$dm)


# a1c values among diagnosed and treated with insulin
a1cdist_dx_rxins = fitdist(nhanescomp$LBXGH[nhanescomp$DIQ050==1], "lnorm", method="mme",weights=round(nhanescomp$WTSAF2YR[nhanescomp$DIQ050==1]/3))
set.seed(100)
a1csamp_dx_rxins=rlnorm(1e5,meanlog=a1cdist_dx_rxins$estimate[1],sdlog = a1cdist_dx_rxins$estimate[2])
quantile(a1csamp_dx_rxins,c(.025,.5,.975))
mean(a1csamp_dx_rxins)

# a1c values among diagnosed and treated with only orals
a1cdist_dx_rxoral = fitdist(as.numeric(na.omit(nhanescomp$LBXGH[(nhanescomp$DIQ070==1)&(nhanescomp$DIQ050==2)])), "lnorm", method="mme",weights=as.numeric(na.omit(round(nhanescomp$WTSAF2YR[(nhanescomp$DIQ070==1)&(nhanescomp$DIQ050==2)]/3))))
set.seed(100)
a1csamp_dx_rxoral=rlnorm(1e5,meanlog=a1cdist_dx_rxoral$estimate[1],sdlog = a1cdist_dx_rxoral$estimate[2])
quantile(a1csamp_dx_rxoral,c(.025,.5,.975))
mean(a1csamp_dx_rxoral)

# a1c values among diagnosed and untreated
a1cdist_dx_norx = fitdist(as.numeric(na.omit(nhanescomp$LBXGH[(nhanescomp$DIQ070==2)&(nhanescomp$DIQ050==2)])), "lnorm", method="mme",weights=as.numeric(na.omit(round(nhanescomp$WTSAF2YR[(nhanescomp$DIQ070==2)&(nhanescomp$DIQ050==2)]/3))))
set.seed(100)
a1csamp_dx_norx=rlnorm(1e5,meanlog=a1cdist_dx_norx$estimate[1],sdlog = a1cdist_dx_norx$estimate[2])
quantile(a1csamp_dx_norx,c(.025,.5,.975))
mean(a1csamp_dx_norx)

# a1c values among undiagnosed 
a1cdist_nodx = fitdist(as.numeric(na.omit(nhanescomp$LBXGH[(nhanescomp$DIQ010==2)])), "lnorm", method="mme",weights=as.numeric(na.omit(round(nhanescomp$WTSAF2YR[(nhanescomp$DIQ010==2)]/3))))
set.seed(100)
a1csamp_nodx=rlnorm(1e5,meanlog=a1cdist_nodx$estimate[1],sdlog = a1cdist_nodx$estimate[2])
quantile(a1csamp_nodx,c(.025,.5,.975))
mean(a1csamp_nodx)
```


# Demographic change approach: 

Demographic change approach: multiply number of people with diagnosed diabetes from 2020-2030 (from IDF estimates) by proportion of people anticipated to be treated with insulin given current estimates (from NHANES) by amount of insulin per person (given distribution of body weights in NHANES).


```{r approach 1}


# https://diabetesatlas.org/data/en/country/211/us.html
undx = .381
total_dm_19 = 30987.90*1000
total_dm_30 = 34374.20*1000

type2prop = 0.93
type2proplo = 0.9
type2prophi = 0.95

proponins1 = sum(nhanescomp$DIQ050==1)/length(nhanescomp$DIQ050)

total_dm = total_dm_19 + (total_dm_30-total_dm_19)/(30-19)*(1:11)

# number of people with diagnosed type 2 dm
  n_dm1=(1-undx)*total_dm*type2prop
  n_dmlo1=(1-undx)*total_dm*type2proplo
  n_dmhi1=(1-undx)*total_dm*type2prophi

# units per person on insulin per day
# The body weight-based dosing estimated that 75% of those treated with insulin 
# require basal insulin only at a dosage of 0.4 IU/kg/day, 
# while the remaining individuals require multiple dose injection therapy 
# totaling 0.6 IU/kg/day.

unitsperday_1 = (0.75*0.4+0.25*0.2)*mean(na.omit(nhanescomp$BMXWT[nhanescomp$DIQ050==1]))*proponins1
unitsperdaylo_1 = (0.75*0.4+0.25*0.2)*quantile(na.omit(nhanescomp$BMXWT[nhanescomp$DIQ050==1]),c(.025))*proponins1
unitsperdayhi_1 = (0.75*0.4+0.25*0.2)*quantile(na.omit(nhanescomp$BMXWT[nhanescomp$DIQ050==1]),c(.975))*proponins1


# total units across entire population of people with diagnosed type 2 dm per year
  vialsperyear_1=unitsperday_1*n_dm1*365.25/1000
  vialsperyearlo_1=unitsperdaylo_1*n_dmlo1*365.25/1000
  vialsperyearhi_1=unitsperdayhi_1*n_dmhi1*365.25/1000


```



# Demographic change plus improved insulin access: 
calculate additional number of people with diabetes not yet on insulin who would be diagnosed if US reached highest OECD country diagnosis level (UK, at 81.5% diagnosis per IDF Diabetes Atlas 9th edition). Estimate insulin necessary given current ADA-EASD guidelines (i.e. start with metformin, then either metformin+ SU, +TZD, +DPP-4, +SGLT-2, or +GLP-1RA). The base case should target HbA1c of 7%. [Sensitivity analyses not yet done: (i) For increased age or significant comorbidities, loosen target to 8%. (ii) try three orals before insulin. ]

```{r approach 2}

#a1c values among those not on insulin, if maximally treated on orals
#titrate those on orals to max dose with effect of -2.5%/2
#those not on orals or undiagnosed need to be with effect of -2.5% 
a1c_after_orals = nhanescomp$LBXGH
a1c_after_orals[is.na(a1c_after_orals)] = mean(a1c_after_orals,na.omit=T)
nhanescomp$DIQ070[is.na(nhanescomp$DIQ070)] = 9
nhanescomp$DIQ050[is.na(nhanescomp$DIQ050)] = 9
nhanescomp$DIQ010[is.na(nhanescomp$DIQ010)] = 9

a1c_after_orals[((nhanescomp$DIQ070==1)&(nhanescomp$DIQ050==2))] = a1c_after_orals[((nhanescomp$DIQ070==1)&(nhanescomp$DIQ050==2))] - 2.5/2
a1c_after_orals[(((nhanescomp$DIQ070==2)&(nhanescomp$DIQ050==2) | (nhanescomp$DIQ010==2)))] = a1c_after_orals[(((nhanescomp$DIQ070==2)&(nhanescomp$DIQ050==2) | (nhanescomp$DIQ010==2)))] - 2.5


# total proportion of people who would now be treated with insulin = 
# proportion already on insulin + 81.5% of proportion w/ a1c >7% after oral therapy

proponins2 = sum(nhanescomp$DIQ050==1)/length(nhanescomp$DIQ050) + 0.815*(table(a1c_after_orals>7)/length(a1c_after_orals))[2]



  n_dm2=total_dm*type2prop
  n_dmlo2=total_dm*type2proplo
  n_dmhi2=total_dm*type2prophi


# units per person on insulin per day
# The body weight-based dosing estimated that 75% of those treated with insulin 
# require basal insulin only at a dosage of 0.4 IU/kg/day, 
# while the remaining individuals require multiple dose injection therapy 
# totaling 0.6 IU/kg/day.

unitsperday_2 = (0.75*0.4+0.25*0.2)*mean(na.omit(nhanescomp$BMXWT[nhanescomp$DIQ050==1]))*proponins2
unitsperdaylo_2 = (0.75*0.4+0.25*0.2)*quantile(na.omit(nhanescomp$BMXWT[nhanescomp$DIQ050==1]),c(.025))*proponins2
unitsperdayhi_2 = (0.75*0.4+0.25*0.2)*quantile(na.omit(nhanescomp$BMXWT[nhanescomp$DIQ050==1]),c(.975))*proponins2

# total units across entire population per year
  vialsperyear_2=unitsperday_2*n_dm2*365.25/1000
  vialsperyearlo_2=unitsperdaylo_2*n_dmlo2*365.25/1000
  vialsperyearhi_2=unitsperdayhi_2*n_dmhi2*365.25/1000



```


#  Match NHANES People to MEPS to get imputed insulin dispensed

```{r match}

meps <- read_csv("mepsdiab021220.csv")

meps = meps[meps$cover_whole_yr==1,]

# meps$days = meps$TOTDAYS
# meps$days[is.na(meps$days)] = 1
meps$units = meps$TOTDOSE
# meps$units[is.na(meps$units)] = 0
meps$unitsperday = (meps$units/365)
meps$unitsperday[meps$unitsperday<=0] = 'NA'
meps$unitsperday[meps$unitsperday=='Inf'] = 'NA'

# meps$DIABAGED[meps$DIABAGED<0]=0

mat_nhanes = nhanescomp %>%
  mutate(baseline_age = RIDAGEYR,
         female = (RIAGENDR==2),
         black = (RIDRETH1==4),
         hisp = (RIDRETH1<3),
         tob = (SMQ040<3),
         sbp = BPXSY1,
         bprx = (BPQ050A==1),
        oraldmrx = (DIQ070==1),
        anti_coag = 0,
        cvd_hx_baseline = ((MCQ160E==1)|(MCQ160F==1)),
        hba1c = LBXGH,
        chol = LBXTC,
        hdl = LBDHDD,
        screat = LBXSCR,
        uacr = URDACT,
        alt = LBXSATSI,
        wt = BMXWT,
        yrsdiab =  (baseline_age - DID040),
        unitsperday = 0) %>%
  dplyr::select(baseline_age,
                female,
                black,
                hisp,
                tob,
                sbp,
                bprx,
                oraldmrx,
                anti_coag,
                cvd_hx_baseline,
                hba1c,
                chol,
                hdl,
                screat,
                uacr,
                alt,
                wt,
                yrsdiab,
                unitsperday)

mat_nhanes$yrsdiab[mat_nhanes$yrsdiab<0] = 'NA'
mat_nhanes$yrsdiab = as.numeric(mat_nhanes$yrsdiab)
mat_nhanes$group = 1
mat_nhanes_imp = mice(mat_nhanes,m=1,maxit=10,meth='pmm',seed=1)
mat_nhanes_comp = complete(mat_nhanes_imp)

meps$DIABAGED = as.numeric(meps$DIABAGED)
meps$DIABAGED[meps$DIABAGED<0] = 0

mat_meps = meps %>%
  mutate(baseline_age = age,
         female = (SEX==2),
         black = (racex==2),
         hisp = 0,
         tob = 0,
         sbp =0,
         bprx = 0,
          oraldmrx = 0,
          anti_coag = 0,
          cvd_hx_baseline = 0,
          hba1c = 0,
          chol = 0, 
          hdl = 0,
          screat = 0,
          uacr = 0,
          alt = 0,
          wt = 0,
         yrsdiab = age - DIABAGED) %>%
  dplyr::select(baseline_age,
                female,
                black,
                hisp,
                tob,
                sbp,
                bprx,
                oraldmrx,
                anti_coag,
                cvd_hx_baseline,
                hba1c,
                chol,
                hdl,
                screat,
                uacr,
                alt,
                wt,
                yrsdiab,
                unitsperday)
mat_meps$yrsdiab[mat_meps$yrsdiab<0]=0
mat_meps$group = 0

mat_meps$unitsperday = as.numeric(mat_meps$unitsperday)

mat_meps_imp = mice(mat_meps,m=1,maxit=10,meth='pmm',seed=1)
mat_meps_comp = complete(mat_meps_imp)

df = rbind(mat_nhanes_comp,mat_meps_comp)

set.seed(1)
match.it <- matchit(group ~ baseline_age + female + black + yrsdiab, data = df, method="nearest", replace = T, caliper = 0.05, ratio = 2)
a <- summary(match.it)
matcheddf <- match.data(match.it)

table1 <- CreateTableOne(vars = c('baseline_age', 'female', 'black', 'yrsdiab'), 
                         data = df, 
                         factorVars = c('female','black'), 
                         strata = 'group')
table1 <- print(table1, 
                printToggle = FALSE, 
                noSpaces = TRUE, smd = TRUE)
knitr::kable(table1[,c(1,2,3,5)],  
      align = 'c', 
      caption = 'Table 1: Comparison of unmatched samples')

knitr::kable(a$nn, digits = 2, align = 'c', 
      caption = 'Table 2: Sample sizes')

table_match1 <- CreateTableOne(vars = c('baseline_age', 'female', 'black', 'yrsdiab'),strata = "group",data = matcheddf,test = FALSE)
table3 <- print(table_match1, 
                printToggle = FALSE, 
                noSpaces = TRUE, smd = TRUE)
knitr::kable(table3,  
      align = 'c', 
      caption = 'Table 3: Summary of balance for matched data')



```


#   Calculate difference between estimated insulin needs (2020-2030) and imputed (based on current 2018 utilization).


```{r diff}

pairsdf = cbind(df[row.names(match.it$match.matrix),"wt"], df[match.it$match.matrix,"unitsperday"], df[row.names(match.it$match.matrix),"hba1c"])

pairsdf[,2][pairsdf[,2]>(mean(pairsdf[,2],na.rm=T)+sd(pairsdf[,2],na.rm=T))] = mean(pairsdf[,2])
pairsdf[,1][pairsdf[,1]>(mean(pairsdf[,1],na.rm=T)+sd(pairsdf[,1],na.rm=T))] = mean(pairsdf[,1])
pairsdf[,2][pairsdf[,2]<(mean(pairsdf[,2],na.rm=T)-sd(pairsdf[,2],na.rm=T))] = mean(pairsdf[,2])
pairsdf[,1][pairsdf[,1]<(mean(pairsdf[,1],na.rm=T)-sd(pairsdf[,1],na.rm=T))] = mean(pairsdf[,1])

a1c_after_orals = pairsdf[,3]
a1c_after_orals[is.na(a1c_after_orals)] = mean(a1c_after_orals,na.omit=T)
nhanescomp$DIQ070[is.na(nhanescomp$DIQ070)] = 9
nhanescomp$DIQ050[is.na(nhanescomp$DIQ050)] = 9
nhanescomp$DIQ010[is.na(nhanescomp$DIQ010)] = 9


a1c_after_orals[((nhanescomp$DIQ070==1)&(nhanescomp$DIQ050==2))] = a1c_after_orals[((nhanescomp$DIQ070==1)&(nhanescomp$DIQ050==2))] - 2.5/2
a1c_after_orals[(((nhanescomp$DIQ070==2)&(nhanescomp$DIQ050==2) | (nhanescomp$DIQ010==2)))] = a1c_after_orals[(((nhanescomp$DIQ070==2)&(nhanescomp$DIQ050==2) | (nhanescomp$DIQ010==2)))] - 2.5

wt = pairsdf[,1]
need = 0*(a1c_after_orals<7) + ((0.75*0.4+0.25*0.2)*wt)*(a1c_after_orals>=7)
summary(need)


# needed minus used
used = pairsdf[,2]
summary(used)
diff_need_minus_use = need - used
summary(diff_need_minus_use)

# proportion over versus under treated
sum((diff_need_minus_use<=0),na.rm = T)/length(diff_need_minus_use)
summary(diff_need_minus_use[((diff_need_minus_use<=0))])


sum((diff_need_minus_use>0),na.rm = T)/length(diff_need_minus_use)
summary(diff_need_minus_use[((diff_need_minus_use>0))])


# a1c distribution in each group

summary(a1c_after_orals[((diff_need_minus_use<=0))])

summary(a1c_after_orals[((diff_need_minus_use>0))])

# weight distribution in each group
summary(wt[((diff_need_minus_use<=0))])

summary(wt[((diff_need_minus_use>0))])



```




